---
title: "SAE_Prediction"
output: html_document
date: "2026-05-28"
---

```{r, echo=F}

```

```{r setup, include=FALSE}
# Hide code by default in R Markdown output.
knitr::opts_chunk$set(echo = FALSE)

# Core data manipulation and reshaping packages.
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(data.table)

# File input and output packages.
library(readxl)
library(gdata)
library(writexl)

# Date and clinical analysis packages.
library(lubridate)
library(comorbidity)
library(table1)

# BigQuery and database connection packages.
library(bigrquery)
library(DBI)
library(dbplyr)

# Modeling, reporting, and plot-composition packages.
library(broom)
library(patchwork)
library(mice)
```


```{r,echo=F}

projectid = "mimic-hr"
bigrquery::bq_auth()

sql <- "SELECT * FROM `physionet-data.mimiciv_3_1_derived.gcs` "
bq_data <- bq_project_query(projectid, query = sql)
gcs = bq_table_download(bq_data)


# Load the SAE analysis dataset from the local CSV file.
#data = fread("df2.3.csv")
cohort = fread("df_sepsis3_clean_with_base_first_day_vars.csv")

icd_exclusion = fread("ICD/sae_exclusion_codes_long.csv")

gcs = gcs%>%filter(stay_id%in%cohort$stay_id)
```

## Seperate Normal GCS at first and Abnormal GCS 
```{r,echo=F}
first_gcs = gcs %>%
  arrange(stay_id, charttime) %>%
  group_by(stay_id) %>%
  filter(!is.na(gcs)) %>%
  slice(1) %>%
  ungroup() %>%
  select(stay_id, first_gcs = gcs, first_gcs_time = charttime)

cohort_with_first_gcs = cohort %>%
  inner_join(first_gcs, by = "stay_id")

cohort_gcs_normal = cohort_with_first_gcs %>%
  filter(first_gcs == 15)

cohort_gcs_abnormal = cohort_with_first_gcs %>%
  filter(first_gcs < 15)

sae_from_gcs = cohort_gcs_normal %>%
  select(stay_id, intime, outtime, first_gcs_time) %>%
  mutate(
    intime = as_datetime(intime),
    outtime = as_datetime(outtime),
    first_gcs_time = as_datetime(first_gcs_time),
    gcs_followup_end = pmin(outtime, intime + days(14), na.rm = TRUE)
  ) %>%
  inner_join(
    gcs %>%
      filter(!is.na(gcs), gcs < 15) %>%
      transmute(
        stay_id,
        abnormal_gcs_time = as_datetime(charttime),
        abnormal_gcs = gcs
      ),
    by = "stay_id"
  ) %>%
  filter(
    abnormal_gcs_time > first_gcs_time,
    abnormal_gcs_time <= gcs_followup_end
  ) %>%
  group_by(stay_id) %>%
  summarise(
    SAE = 1,
    SAE_time = min(abnormal_gcs_time),
    .groups = "drop"
  )

cohort_gcs_normal = cohort_gcs_normal %>%
  left_join(sae_from_gcs, by = "stay_id") %>%
  mutate(SAE = replace_na(SAE, 0))

```

## GCS scores up to 14 days for normal first GCS cohort
```{r,echo=F}
cohort_normal_gcs = cohort_gcs_normal

gcs_14d_cohort_normal_gcs = cohort_normal_gcs %>%
  select(stay_id, intime, outtime) %>%
  mutate(
    intime = as_datetime(intime),
    outtime = as_datetime(outtime),
    gcs_followup_end = pmin(outtime, intime + days(14), na.rm = TRUE)
  ) %>%
  inner_join(
    gcs %>%
      filter(!is.na(gcs)) %>%
      transmute(
        stay_id,
        gcs_time = as_datetime(charttime),
        gcs
      ),
    by = "stay_id"
  ) %>%
  filter(
    gcs_time >= intime,
    gcs_time <= gcs_followup_end
  ) %>%
  arrange(stay_id, gcs_time)

```



### Table 1


### Table 2


### Trajectory Model (k=2/k=3)

Using table "gcs_14d_cohort_normal_gcs" as longitudinal data.

Then, create 2 columns indicating the cluster of the patient in table "cohort_gcs_normal"


### Risk Model

Build two logistic regression model of hospital_death_flag ~ cluster + first_day_SOFA + "first_day_invasive_vent"         +"first_day_noninvasive_vent"       
+"first_day_hfnc"                   
+"first_day_tracheostomy"

On k=2 cluster and k=3 cluster separately. Plot forest plot with ggplot2 package. 

### Prediction Model

#### Model-1
Predict hospital mortality using all the first_day data and demographics(gender,age)

Plot the AUROC for three models and SHAP for XGboost


#### Model-2
Predict cluster using first day data (k=2/k=3 separately).

Plot the AUROC for three models and SHAP for XGboost. (There will be two AUROC and two SHAP)
